Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-10-22 GWLDCII.tex
more tweaking to denominator paragraph
changes to intro

In contrast, the first paper in this series focuses on combinatorial and matroidal properties of the Wilson loop diagrams. The main focus of that paper is enumerating the positroid cells that are associated to Wilson loop diagrams with a fixed number of vertices and propagators, as well as counting the number of Wilson loop diagrams that map to the same positroid.

@@

This paper continues this work by identifying which positroid cells are associated to which Wilson loop diagram: specifically, we give an algorithm to read off the Grassmann necklace of the associated positroid cell from the Wilson loop diagram. We also show that the positroid cells associated to Wilson loop diagrams are all $3k$-dimensional, where $k$ is the number of propagators in the diagram; by contrast, the BCFW cells are all $4k$-dimensional \cite{Arkani-Hamed:2013jha}. This apparent contradiction led to the work in \cite{non-orientable}, which showed that the actual geometric space defined by the Wilson loop diagrams is a non-orientable $k$-dimensional bundle over this $3k$ dimensional manifold in $\Gr(k,n)$.

As noted above, the sum of the Feynman integrals associated to Wilson loop diagrams gives the volume of a convex polytope. If the singularities appearing in individual integrals cancel out in the overall sum, then the overall volume is finite. This cancellation has been conjectured in general \cite{Hodges2013} and verified explicitly in a few cases \cite{casestudy, HeslopStewart}, but it is hard to prove in general due to the fact that the relationship between Wilson loop diagrams and $3k$-dimensional positroid cells of $\Gr(k,n)$ is neither one-to-one nor onto. In this paper, we give a characterization of the factors appearing in the denominator of these Feynman integrals in terms of the Grassmann necklace of the associated positroid, thus providing a much simpler and more tractable link between the singularities and the diagrams.

% when computing a finite volume to a geometric object defined by the Wilson loop diagrams, one must ensure that any singularities cancel. This has been verified explicitly in a few cases \cite{casestudy, HeslopStewart}, but it has been hard to prove in general due to the fact that the relationship between Wilson loop diagrams and $3k$ dimensional positroid cells of $\Gr(k,n)$ is neither one-to-one nor onto. In this paper, we give a characterisation of the denominator of the Feynman integral defined by a Wilson loop diagram in terms of the algorithm that associated the same diagram to a Grassman Necklace.

@@

Finally in Section~\ref{sec poles} we characterize the denominator of the integrand of a Wilson loop diagram in terms of how propagators contribute to the Grassmann necklace.  This characterization is given explicitly by Algorithm~\ref{alg WLD to denom via GN} and Theorem~\ref{thm denom}.

@@

This formulation for the denominator comes from the physics literature: see equation (2) in \cite{HeslopStewart}. In Section \ref{sec poles}, we give a geometric meaning to the factors, $\{R_e | e \in V\}$, of $R(W)$.  Given a Grassmann necklace representation of the positroid cell $\Sigma(W)$ ((see Definition~\ref{def:grassmann necklace} and Algorithm \ref{alg:put GN on WLD}), the $R_e$ are exactly the irreducible factors of the minors of $C(W)$ defined by the Grassman necklace. In particular, \emph{which} minors of $C(W)$ contribut to $R(W)$ is geometrically determined. Furthermore, when two Wilson loop diagrams correspond to the same postroid cell, they still have different $R(W)$. In Algorithm~\ref{alg WLD to denom via GN} we give an algorithm to read $R(W)$ directly from the Wilson loop diagram, paving the way for a general study of these poles in future.

We end this section with an example of computing $C(W)$ and $R(W)$ for a particular Wilson loop diagram $W$.

>>>>

In contrast, the first paper in this series focuses on combinatorial and matroidal properties of the Wilson loop diagrams. The main focus of that paper is enumerating the positroid cells that are associated to Wilson loop diagrams with a fixed number of vertices and propagators, as well as counting the number of Wilson loop diagrams that map to the same positroid. The results of the first paper are not needed for this paper.

@@

This paper continues this body of work by establishing concrete connections between the combinatorics of the Wilson loop diagram as determined by the physics literature, and the combinatorics of the assoicated positroid cell (as identified in \cite{wilsonloop}). For instance, a Grassman Necklace is one of the many equivalent combinatorial objects uniquely identifying positroid cells. In this paper, we give an algorithm to read off the Grassmann necklace of the associated positroid cell from the Wilson loop diagram. We also show that the positroid cells associated to Wilson loop diagrams are all $3k$-dimensional, where $k$ is the number of propagators in the diagram; by contrast, the BCFW cells are all $4k$-dimensional \cite{Arkani-Hamed:2013jha}. This apparent contradiction led to the work in \cite{non-orientable}, which showed that the actual geometric space defined by the Wilson loop diagrams is a non-orientable $k$-dimensional bundle over this $3k$ dimensional manifold in $\Gr(k,n)$.

SYM $N=4$ is a finite theory \cite{Adamo:2011pv,Adamo:2012xe,Arkani-Hamed:2013jha}. That is, the scattering amplitudes for particle interactions are finite. Since the (sum of) integrals associated to these scattering amplitudes can be associated to (the union of) convex polytopes, as the  noted above, one considers the sum of integrals as defining a volume to said union of polytopes. However, each individual integral has singularities occuring on the boundaries of the corresponding convex polytope. If the singularities cancel out in the overall sum, then the overall volume is finite. This cancellation has been conjectured in general \cite{Hodges2013} and verified explicitly in a few cases \cite{casestudy, HeslopStewart}, but it is hard to prove in general. In this paper, we give a goemetric characterization of the factors appearing in the denominator of these integrals.  The set of irreducible factors in the denominator, as defined in the physics literature \cite{HeslopStewart}, are exactly the irreducible factors of the minors defined by the Grassmann necklace in a physically significant representation of the positroid cell assoicated to a diagram. In other words, the singluarities arising from tree level particle interactions in SYM $N=4$ theory do not just come from the Feynman rules describing the theory. They can also be seen directly from the associated geometry, given the physically correct representation.  We hope these results will pave the way for a general study of these poles in future.

%This paper continues this work by identifying which positroid cells are associated to which Wilson loop diagram: specifically, we give an algorithm to read off the Grassmann necklace of the associated positroid cell from the Wilson loop diagram. We also show that the positroid cells associated to Wilson loop diagrams are all $3k$-dimensional, where $k$ is the number of propagators in the diagram; by contrast, the BCFW cells are all $4k$-dimensional \cite{Arkani-Hamed:2013jha}. This apparent contradiction led to the work in \cite{non-orientable}, which showed that the actual geometric space defined by the Wilson loop diagrams is a non-orientable $k$-dimensional bundle over this $3k$ dimensional manifold in $\Gr(k,n)$.

%As noted above, the sum of the Feynman integrals associated to Wilson loop diagrams gives the volume of a convex polytope. If the singularities appearing in individual integrals cancel out in the overall sum, then the overall volume is finite. This cancellation has been conjectured in general \cite{Hodges2013} and verified explicitly in a few cases \cite{casestudy, HeslopStewart}, but it is hard to prove in general due to the fact that the relationship between Wilson loop diagrams and $3k$-dimensional positroid cells of $\Gr(k,n)$ is neither one-to-one nor onto. In this paper, we give a characterization of the factors appearing in the denominator of these Feynman integrals in terms of the Grassmann necklace of the associated positroid, thus providing a much simpler and more tractable link between the singularities and the diagrams.

@@

Finally in Section~\ref{sec poles} we characterize the denominator of the integrand of a Wilson loop diagram in terms the associated matrix and the Grassmann necklace. This characterization is given explicitly by Algorithm~\ref{alg WLD to denom via GN} and Theorem~\ref{thm denom}. Given the physically correct representation of the positroid cell assoicated to a Wilson loop diagram, the irreducible factors of the denominator are exactly the set of irreducible factors of the minors of said representation given by the Grassmann necklace.

@@

This formulation for the denominator comes from the physics literature: see equation (2) in \cite{HeslopStewart}. In Section \ref{sec poles}, we give a geometric meaning to the factors, $\{R_e | e \in V\}$, of $R(W)$.  Given a Grassmann necklace representation of the positroid cell $\Sigma(W)$ ((see Definition~\ref{def:grassmann necklace} and Algorithm \ref{alg:put GN on WLD}), the $R_e$ are exactly the irreducible factors of the minors of $C(W)$ defined by the Grassman necklace. In particular, \emph{which} minors of $C(W)$ contribut to $R(W)$ is geometrically determined. 
We end this section with an example of computing $C(W)$ and $R(W)$ for a particular Wilson loop diagram $W$.
